\documentclass[12pt]{article}

\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{thmtools}
\usepackage{amssymb}
\usepackage{enumitem}
\usepackage{float}
\usepackage{gensymb}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{mathtools}
\usepackage{multirow}
\usepackage{physics}
\usepackage{tabularx}
\usepackage{wrapfig}
\usepackage[utf8]{inputenc}


\setlist{parsep=1em}

\newcommand{\succr}[1]{#1 \textrm{++}}
\newcommand{\minus}{\textrm{ --- }}

\newenvironment{df}[2][\relax]{
\vspace{4mm}
    \noindent\setlength{\parskip}{1em}%
    \ifx#1\relax%
        \textbf{#2. }%
    \else%
        \textbf{#2} (#1)\textbf{. }%
    \fi%
    \ignorespaces%
}{}

\theoremstyle{remark}
\newtheorem*{remark}{Remark}


\title{Definitions in Analysis I by Terence Tao}
\author{Li Thon Hong}
\date{}

\begin{document}

\maketitle
\tableofcontents
\vspace{5mm}
\begin{remark}
    This document follows the third edition of \textit{Analysis I}.
\end{remark}

\pagebreak


\section{Introduction}
No definitions found in this chapter.

\pagebreak


\section{Starting at the beginning: the natural numbers}
\subsection{The Peano axioms}

\begin{df}{Axiom 2.1}
    $0$ is a natural number.
\end{df}


\begin{df}{Axiom 2.2}
    If $n$ is a natural number, then $\succr{n}$ is also a natural number.
\end{df}


\begin{df}{Definition 2.1.3}
    We define $1$ to be the number $\succr{0}$, $2$ to be the number $\succr{(\succr{0})}$, 3 to be the number $\succr{(\succr{(\succr{0})})}$, etc.
    In other words, $1 \coloneqq \succr{0}$, $2 \coloneqq \succr{1}$, $3 \coloneqq \succr{2}$, etc.
\end{df}


\begin{df}{Axiom 2.3}
    $0$ is not the successor of any natural number, i.e., we have $\succr{n} \ne 0$ for every natural number $n$.
\end{df}


\begin{df}{Axiom 2.4}
    Different natural numbers must have different successors; i.e. if $n$, $m$ are natural numbers and $n \ne m$, then $\succr{n} = \succr{m}$.
    Equivalently, if $\succr{n} = \succr{m}$, then we must have $n = m$.    
\end{df}


\begin{df}[Principle of mathematical induction]{Axiom 2.5}
    Let $P(n)$ be any property pertaining to a natural number $n$.
    Suppose that $P(0)$ is true, and suppose that whenever $P(n)$ is true, $P(\succr{n})$ is also true.
    Then $P(n)$ is true for every natural number $n$.
\end{df}


\subsection{Addition}

\begin{df}[Addition of natural numbers]{Definition 2.2.1}
    Let $m$ be a natural number.
    To add $0$ to $m$, we define $0 + m \coloneqq m$.
    Now, suppose inductively that we have defined how to add $n$ to $m$.
    Then, we can add $\succr{n}$ to m by defining $(\succr{n}) + m \coloneqq \succr{(n + m)}$.%
\end{df}


\begin{df}{Lemma 2.2.2}
    For any natural number $n$, $n + 0 = n$.
\end{df}


\begin{df}{Lemma 2.2.3}
    For any natural numbers $n$ and $m$, $n + (\succr{m}) = \succr{(n + m)}$.
\end{df}


\begin{df}[Addition is commutative]{Proposition 2.2.4}
    For any natural numbers $n$ and $m$, $n + m = m + n$.
\end{df}


\begin{df}[Addition is associative]{Proposition 2.2.5}
    For any natural numbers $a$, $b$, $c$, we have $(a + b) + c = a +(b + c)$.
\end{df}


\begin{df}[Cancellation law]{Proposition 2.2.6}
    Let $a$, $b$, $c$ be natural numbers such that $a + b = a + c$.
    Then we have $b = c$.
\end{df}


\begin{df}[Positive natural numbers]{Definition 2.2.7}
    A natural number $b$ is said to be \emph{positive} iff it is not equal to $0$.    
\end{df}


\begin{df}{Proposition 2.2.8}
    If $a$ is positive and $b$ is a natural number, then $a + b$ is positive.
\end{df}


\begin{df}{Corollary 2.2.9}
    If $a$ and $b$ are natural numbers such that $a + b = 0$, then $a = 0$ and $b = 0$.
\end{df}


\begin{df}{Lemma 2.2.10}
    Let $a$ be a positive number.
    Then there exists exactly one natural number $b$ such that $\succr{b} = a$.
\end{df}


\begin{df}[Ordering of the natural numbers]{Definition 2.2.11}
    Let $n$ and $m$ be natural numbers.
    We say that $n$ is greater than or equal to $m$, and write $n \ge m$ or $m \le n$, iff we have $n = m + a$ for some natural number $a$.
    We say that $n$ is \emph{strictly greater than} $m$, and write $n > m$ or $m < n$, iff $n \ge m$ and $n \ne m$.
\end{df}


\begin{df}[Basic properties of order for natural numbers]{Proposition 2.2.12}
    Let $a$, $b$, $c$ be natural numbers. Then
    \begin{enumerate}[topsep=0mm,parsep=0mm,label=(\alph*)]
        \item \textit{Reflexivity: $a \ge a$.}
        \item \textit{Transitivity: If  $a \ge b$ and $b \ge c$, then $a \ge c$.}
        \item \textit{Anti-symmetry: If $a \ge b$ and $b \ge a$, then $a = b$.}
        \item \textit{Preservation of order: $a \ge b$ if and only if $a + c \ge b + c$.}
        \item \textit{$a < b$ if and only if $\succr{a} < b$.}
        \item \textit{$a < b$ if and only if $b = a + d$ for some positive number $d$.}
    \end{enumerate}    
\end{df}


\begin{df}[Trichotomy of order for natural number]{Proposition 2.2.13}
    Let $a$ and $b$ be natural numbers.
    Then exactly one of the following statements is true: $a < b$, $a = b$, or $a > b$.
\end{df}


\begin{df}[Strong principle of induction]{Proposition 2.2.14}
    Let $m_0$ be a natural number, and let $P(m)$ be a property pertaining to an arbitrary natural number $m$.
    Suppose that for each $m \ge m_0$, we have the following implication: if $P(m')$ is true for all natural numbers $m_0 \le m' < m$, then $P(m)$ is also true.
    Then we can conclude that $P(m)$ is true for all natural numbers $m \ge m_0$.
\end{df}


\subsection{Multiplication}

\begin{df}[Multiplication of natural numbers]{Definition 2.3.1}
    Let $m$ be a natural number.
    To multiply $0$ to $m$, we define $0 \times m \coloneqq 0$.
    Now, suppose inductively that we have defined how to multiply $n$ to $m$.
    Then, we can multiply $\succr{n}$ to m by defining $(\succr{n}) \times m \coloneqq (n \times m) + m$.%
\end{df}


\begin{df}[Multiplication is commutative]{Lemma 2.3.2}
    Let $n$, $m$ be natural numbers.
    Then $n \times m = m \times n$.    
\end{df}


\begin{df}[Positive natural numbers have no zero divisors]{Lemma 2.3.3}
    Let $n$, $m$ be natural numbers.
    Then $n \times m = 0$ if and only if at least one of $n$, $m$ is equal to $0$.
    In particular, if $n$ and $m$ are both positive, then $nm$ is also positive.
\end{df}


\begin{df}[Distributive law]{Proposition 2.3.4}
    For any natural numbers $a$, $b$, $c$, we have $a(b + c) = ab + ac$ and $(b + c)a = ba + ca$.
\end{df}


\begin{df}[Multiplication is associative]{Proposition 2.3.5}
    For any natural numbers $a$, $b$, $c$, we have $(a \times b)\times c = a \times (b \times c)$
\end{df}


\begin{df}[Multiplication preserves order]{Proposition 2.3.6}
    If $a$, $b$ are natural numbers such that $a < b$, and $c$ is positive, then $ac < bc$.
\end{df}


\begin{df}[Cancellation law]{Corollary 2.3.7}
    Let $a$, $b$, $c$ be natural numbers such that $ac = bc$ and $c$ is nonzero.
    Then $a = b$.
\end{df}


\begin{df}[Euclidean algorithm]{Proposition 2.3.9}
    Let $n$ be a natural number, and let $q$ be a positive number.
    Then there exist natural numbers $m$, $r$ such that $0 \le r < q$ and $n = mq + r$.
\end{df}


\begin{df}[Exponentiation for natural numbers]{Definition 2.3.11}
    Let $m$ be a natural number.
    To raise $m$ to the power $0$, we define $m^0 \coloneqq 1$; in particular, we define $0^0 \coloneqq 1$.
    Now suppose recursively that $m^n$ has been defined for some natural number $n$, then we define $m^{\succr{n}} \coloneqq m^n \times m$.
\end{df}


\pagebreak


\section{Set theory}
\subsection{Fundamentals}
\begin{df}[Sets are objects]{Axiom 3.1}
    If $A$ is a set, then $A$ is also an object.
    In particular, given two sets $A$ and $B$, it is meaningful to ask whether $A$ is also an element of $B$.
\end{df}


\begin{df}[Equality of sets]{Definition 3.1.4}
    Two sets $A$ and $B$ are \emph{equal}, $A = B$, iff every element of $A$ is an element of $B$ and vice versa.
    To put it another way, $A = B$ if and only if every element $x$ of $A$ belongs also to $B$, and every element $y$ of $B$ belongs also to $A$.
\end{df}


\begin{df}[Empty set]{Axiom 3.2}
    There exists a set $\varnothing$, known as the \emph{empty set}, which contains no elements, i.e., for every object $x$ we have $x \notin \varnothing$.
\end{df}


\begin{df}[Single choice]{Lemma 3.1.6}
    Let $A$ be a non-empty set.
    Then there exists an object $x$ such that $x \in A$.
\end{df}


\begin{df}[Singleton sets and pair sets]{Axiom 3.3}
    If $a$ is an object, then there exists a set $\{a\}$ whose only element is $a$, i.e., for every object $y$, we have $y \in \{ a \}$ if and only if $y = a$; we refer to $\{ a \}$ as the singleton set whose element is $a$.
    Furthermore, if $a$ and $b$ are objects, then there exists a set $\{a, b\}$ whose only elements are $a$ and $b$; i.e., for every object $y$, we have $y \in \{a, b\}$ if and only if $y = a$ or $y = b$; we refer to this set as the pair set formed by $a$ and $b$.
\end{df}


\begin{df}[Pairwise union]{Axiom 3.4}
    Given any two sets $A$, $B$, there exists a set $A \cup B$, called the \emph{union} $A \cup B$ of $A$ and $B$, whose elements consist of all the elements which belong to $A$, or $B$ or both.
    In other words, for any object $x$,
    {\setlength{\belowdisplayskip}{-10pt} \setlength{\belowdisplayshortskip}{-10pt}
    \[
        x \in A \cup B \iff (x \in A \text{ or } x \in B).
    \]
    }
\end{df}


\begin{df}{Lemma 3.1.13}
    If $a$ and $b$ are objects, then $\{ a, b \} = \{ a \} \cup \{ b \}$.
    If $A$, $B$, $C$ are sets, then the union operation is commutative (i.e., $A \cup B = B \cup A$) and associative (i.e., $(A \cup B) \cup C = A \cup (B \cup C)$).
    Also, we have $A \cup A = A \cup \varnothing = \varnothing \cup A = A$.
\end{df}


\begin{df}[Subsets]{Definition 3.1.15}
    Let $A$, $B$ be sets.
    We say that $A$ is a \emph{subset} of $B$, denoted $A \subseteq B$, iff every element of $A$ is also an element of $B$, i.e.
    \[
        \text{For any object } x, \quad x \in a \implies x \in B.
    \]
    We say that $A$ is a \emph{proper subset} of $B$, denoted $A \subsetneq B$, if $A \subseteq B$ and $A \ne B$.
\end{df}


\begin{df}[Sets are partially ordered by set inclusion]{Proposition 3.1.18}
    Let $A$, $B$, $C$ be sets.
    If $A \subseteq B$ and $B \subseteq C$, then $A \subseteq C$.
    If $A \subseteq B$ and $B \subseteq A$, then $A = B$.
    Finally, if $A \subsetneq B$ and $B \subsetneq C$, then $A \subsetneq C$.
\end{df}


\begin{df}[Axiom of specification/axiom of separation]{Axiom 3.5}
    Let $A$ be a set, and for each $x \in A$, let $P(x)$ be a property pertaining to $x$ (i.e., $P(x)$ is either a true statement or a false statement).
    Then there exists a set, called $\{x \in A : P(x) \text{ is true} \}$ (or simply $\{x \in A : P(x)\}$ for short), whose elements are precisely the elements $x$ in $A$ for which $P(x)$ is true.
    In other words, for any object $y$,
    {\setlength{\belowdisplayskip}{-10pt} \setlength{\belowdisplayshortskip}{-10pt}
    \[
        y \in \{ x \in A : P(x) \text{ is true} \} \iff (y \in A \text{ and } P(y) \text{ is true}).
    \]
    }
\end{df}


\begin{df}[Intersections]{Definition 3.1.23}
    The \emph{intersection} $S_1 \cap S_2$ of two sets is defined to be the set
    {\setlength{\belowdisplayskip}{-10pt} \setlength{\belowdisplayshortskip}{-10pt}
    \[
        S_1 \cap S_2 \coloneqq \{ x \in S_1 : x \in S_2 \}.
    \]
    }
\end{df}

\begin{df}[Difference set]{Definition 3.1.27}
    Given two sets $A$ and $B$, we define the set $A - B$ or $A \setminus B$ to be the set $A$ with any elements of $B$ removed:
    {\setlength{\belowdisplayskip}{-10pt} \setlength{\belowdisplayshortskip}{-10pt}
    \[
        A \setminus B \coloneqq \{ x \in A : x \notin B \}.
    \]
    }
\end{df}


\begin{df}[Sets form a boolean algebra]{Proposition 3.1.28}
    Let $A, B, C$ be sets, and let $X$ be a set containing $A, B, C$ as subsets.
    \begin{enumerate}[topsep=0mm,parsep=0mm,label=(\alph*)]
        \item (Minimal element) We have $A \cup \varnothing = A$ and $A \cap \varnothing = \varnothing$.
        \item (Maximal element) We have $A \cup X = X$ and $A \cap X = A$.
        \item (Identitiy) We have $A \cap A= A$ and $A \cup A = A$.
        \item (Commutativity) We have $A \cup B = B \cup A$ and $A \cap B = B \cap A$.
        \item (Associativity) We have $(A \cup B) \cup C = A \cup (B \cup C)$and $(A \cap B) \cap C = A \cap (B \cap C)$.
        \item (Distributivity) We have $A \cap (B \cup C) = (A \cap B) \cup (A \cap C)$ and $A \cup (B \cap C)=(A \cup B) \cap (A \cup C)$.
        \item (Partition) We have $A \cup (X \setminus A) = X$ and $A \cap (X \setminus A) = \varnothing$.
        \item (De Morgan Laws) We have $X \setminus (A \cup B) = (X \setminus A) \cap (X \setminus B)$ and $X \setminus (A \cap B) = (X \setminus A) \cup (X \setminus B)$.
    \end{enumerate}
\end{df}


\begin{df}[Replacement]{Axiom 3.6}
    Let $A$ be a set.
    For any object $x \in A$ and any object $y$, suppose we have a statement $P(x, y)$ pertaining to $x$ and $y$, such that for each $x \in A$ there is at most one $y$ for which $P(x, y)$ is true.
    Then there exists a set $\{ y : P(x, y) \text{ is true for some } x \in A \}$, such that for any object $z$,
    {\setlength{\belowdisplayskip}{-10pt} \setlength{\belowdisplayshortskip}{-10pt}
    \[
        z \in \{ y : P(x, y) \text{ is true for some } x \in A \} \iff P(x, z) \text{ is true for some } x \in A.
    \]
    }
\end{df}


\begin{df}[Infinity]{Axiom 3.7}
    There exists a set $\mathbb{N}$, whose elements are called natural numbers, as well as an object $0$ in $\mathbb{N}$, and an object $\succr{n}$ assigned to every natural number $n \in \mathbb{N}$, such that the Peano axioms (Axioms 2.1 - 2.5) hold.
\end{df}


\subsection{Russell's paradox}
\begin{df}[Universal specification]{Axiom 3.8}
    (\textit{Dangerous!})
    Suppose for every object $x$ we have a property $P(x)$ pertaining to $x$ (so that for every $x$, $P(x)$ is either a true statement or a false statement).
    Then there exists a set $\{x : P(x) \text{ is true} \}$ such that for every object $y$,
    {\setlength{\belowdisplayskip}{-10pt} \setlength{\belowdisplayshortskip}{-10pt}
    \[
        y \in \{x : P(x) \text{ is true} \} \iff P(y) \text{ is true.}
    \]
    }
\end{df}


\begin{df}[Regularity]{Axiom 3.9}
    If $A$ is a non-empty set, then there is at least one element $x$ of $A$ which is either not a set, or is disjoint from $A$.
\end{df}


\subsection{Functions}
\begin{df}[Functions]{Definition 3.3.1}
     Let $X$, $Y$ be sets, and let $P(x, y)$ be a property pertaining to an object $x \in X$ and an object $y \in Y$, such that for every $x \in X$, there is exactly one $y \in Y$ for which $P(x, y)$ is true (this is sometimes known as the \emph{vertical line test}).
     Then we define the \emph{function $f : X \to Y$ defined by $P$ on the domain $X$ and range $Y$} to be the object which, given any input $x \in X$, assigns an output $f(x) \in Y$, defined to be the unique object $f(x)$ for which $P(x, f(x))$ is true.
     Thus, for any $x \in X$ and $y \in Y$,
     {\setlength{\belowdisplayskip}{-10pt} \setlength{\belowdisplayshortskip}{-10pt}
     \[
        y = f(x) \iff P(x, y) \text{ is true.}
    \]
    }
\end{df}


\begin{df}[Equality of functions]{Definition 3.3.7}
    Two functions $f : X \to Y$, $g : X \to Y$ with the same domain and range are said to be equal, $f = g$, if and only if $f(x) = g(x)$ for all $x \in X$.
    (If $f(x)$ and $g(x)$ agree for some values of $x$, but not others, then we do not consider $f$ and $g$ to be equal.)
\end{df}


\begin{df}[Composition]{Definition 3.3.10}
     Let $f : X \to Y$ and $g : Y \to Z$ be two functions, such that the range of $f$ is the same set as the domain of $g$.
     We then define the composition $g \circ f : X \to Z$ of the two functions $g$ and $f$ to be the function defined explicitly by the formula
     \[
        (g \circ f)(x) \coloneqq g(f(x)).
    \]
    If the range of $f$ does not match the domain of $g$, we leave the composition $g \circ f$ undefined.
\end{df}


\begin{df}[Composition is associative]{Lemma 3.1.12}
    Let $f : Z \to W$, $g : Y \to Z$, and $h : X \to Y$ be functions. Then $f \circ (g \circ h) = (f \circ g) \circ h$.
\end{df}


\begin{df}[One-to-one function]{Definition 3.3.14}
    A function $f$ is \emph{one-to-one} (or \emph{injective}) if different elements map to different elements:
    \[
        x \ne x' \implies f(x) \ne f(x').
    \]
    Equivalently, a function is one-to-one if
    {\setlength{\belowdisplayskip}{-10pt} \setlength{\belowdisplayshortskip}{-10pt}
    \[
        f(x) = f(x') \implies x = x'.
    \]
    }
\end{df}


\begin{df}[Onto functions]{Definition 3.3.17}
    A function $f$ is \emph{onto} (or \emph{surjective}) if $f(X) = Y$, i.e., every element in $Y$ comes from applying $f$ to some element in $X$:
    {\setlength{\belowdisplayskip}{-10pt} \setlength{\belowdisplayshortskip}{-10pt}
    \[
        \text{For every } y \in Y\text{, there exists } x \in X \text{ such that } f(x) = y.
    \]
    }
\end{df}


\begin{df}[Bijective functions]{Definition 3.3.20}
    Functions $f : X \to Y$ which are both one-to-one and onto are also called \emph{bijective} or \emph{invertible}.
\end{df}


\subsection{Images and inverse images}
\begin{df}[Images of sets]{Definition 3.4.1}
    If $f : X \to Y$ is a function from $X$ to $Y$, and $S$ is a set in $X$, we define $f(S)$ to be the set
    \[
        f(S) \coloneqq \{ f(x): x \in S \};
    \]
    this set is a subset of $Y$, and is sometimes called the \emph{image} of $S$ under the map $f$.
    In other words,
    \[
        x \in S \implies f(x) \in f(S),
    \]
    and
    \[
        y \in f(S) \iff y = f(x) \text{ for some } x \in S.
    \]
\end{df}


\begin{df}[Inverse images]{Definition 3.4.4}
     If $U$ is a subset of $Y$, we define the set $f^{-1}(U)$ to be the set
     \[
        f^{-1}(U) \coloneqq \{ x \in X : f(x) \in U \}.
    \]
    In other words, $f^{-1}(U)$ consists of all the elements of $X$ which map into $U$:
    \[
        f(x) \in U \iff x \in f^{-1}(U).
    \]
    We call $f^{-1}(U)$ the inverse image of $U$.
\end{df}


\begin{df}[Power set axiom]{Axiom 3.10}
     Let $X$ and $Y$ be sets.
     Then there exists a set, denoted $Y^X$, which consists of all the functions from $X$ to $Y$, thus
     \[
        f \in Y^X \iff (f \text{ is a function with domain } X \text{ and range } Y).
    \]
\end{df}


\begin{df}{Lemma 3.4.9}
    Let $X$ be a set. Then the set
    \[
        \{ Y : Y \text{ is a subset of } X \}
    \]
    is a set.
    This set is known as the \emph{power set} of $X$ and is denoted $2^X$.
\end{df}


\begin{df}[Union]{Axiom 3.11}
    Let $A$ be a set, all of whose elements are themselves sets.
    Then there exists a set $\bigcup A$ whose elements are precisely those objects which are elements of the elements of $A$, thus for all objects $x$
    \[
        x \in \bigcup A \iff (x \in S \text{ for some } S \in A). 
    \]

     If one has some set $I$, and for every element $\alpha \in I$ we have some set $A_\alpha$, then we can form the union set $\bigcup _{\alpha \in I} A_\alpha$ by defining
     \[
        \bigcup_{\alpha \in I} A_\alpha \coloneqq \bigcup \{ A_\alpha : \alpha \in I \}.
     \]
     More generally, for any objects $y$,
     \[
        y \in \bigcup_{\alpha \in I} A_\alpha \iff (y \in A_\alpha \text{ for some } \alpha \in I). \tag{3.2}\label{eq:3.2}
     \]
     We often refer to $I$ as an \emph{index set}, and the elements $\alpha$ of this index set as \emph{labels}; the sets $A_\alpha$ are then called a \emph{family of sets}, and are indexed by the labels $\alpha \in I$.

    Similarly, we can form intersections of families of sets.
    Given any non-empty set $I$, and given an assignment of a set $A_\alpha$ to each $\alpha \in I$, we can define the intersection $\bigcap _{\alpha \in I} A_\alpha$ by first choosing some element $\beta$ of $I$ (which we can do since $I$ is non-empty), and setting
    \[
        \bigcap_{\alpha \in I} A_\alpha \coloneqq \{x \in A_\beta : x \in A_\alpha \text{ for all } \alpha \in I \}. \tag{3.3}\label{eq:3.3}
    \]
    Then, for any object $y$,
    {\setlength{\belowdisplayskip}{-10pt} \setlength{\belowdisplayshortskip}{-10pt}
    \[
        y \in \bigcap_{\alpha \in I} A_\alpha \iff (y \in A_\alpha \text{ for all } \alpha \in I). \tag{3.4}\label{eq:3.4}
    \]
    }
\end{df}


\subsection{Cartesian products}
\begin{df}[Ordered pair]{Definition 3.5.1}
    If $x$ and $y$ are any objects (possibly equal), we define the ordered pair $(x, y)$ to be a new object, consisting of $x$ as its first component and $y$ as its second component.
    Two ordered pairs $(x, y)$ and $(x', y')$ are considered equal if and only if both their components match, i.e.
    {\setlength{\belowdisplayskip}{-10pt} \setlength{\belowdisplayshortskip}{-10pt}
    \begin{equation}
        (x, y) = (x', y') \iff (x = x' \text{ and } y = y'). \tag{3.5}\label{eq:3.5}
    \end{equation}
    }
\end{df}


\begin{df}[Cartesian product]{Definition 3.5.4}
    If $X$ and $Y$ are sets, then we define the Cartesian product $X \times Y$ to be the collection of ordered pairs, whose first component lies in $X$ and second component lies in $Y$, thus
    \[
        X \times Y = \{(x, y): x \in X, y \in Y \}
    \]
    or equivalently
    {\setlength{\belowdisplayskip}{-10pt} \setlength{\belowdisplayshortskip}{-10pt}
    \[
        a \in (X \times Y) \iff (a = (x, y) \text{ for some } x \in X \text{ and } y \in Y ).
    \]
    }
\end{df}


\begin{df}[Ordered $n$-tuple and $n$-fold Cartesian product]{Definition 3.5.7}
    Let $n$ be a natural number.
    An ordered $n$-tuple $(x_i)_{1 \le i \le n}$ (also denoted $(x_1, \ldots , x_n)$) is a collection of objects $x_i$, one for every natural number $i$ between $1$ and $n$;
    we refer to $xi$ as the \emph{$i$\textsuperscript{th} component} of the ordered
    $n$-tuple.
    
    Two ordered $n$-tuples $(x_i)_{1 \le i \le n}$ and $(y_i)_{1 \le i \le n}$ are said to be equal iff $x_i = y_i$ for all $1 \le i \le n$.
    
    If $(X_i)_{1 \le i \le n}$ is an ordered $n$-tuple of sets, we define their \emph{Cartesian product} $\prod_{1 \le i \le n} X_i$ (also denoted $\prod^n_{i=1} X_i$ or $X_1 \times \ldots \times X_n$) by
    \[
        \prod_{1 \le i \le n} X_i \coloneqq \{(x_i)_{1 \le i \le n} : x_i \in X_i \text{ for all } 1 \le i \le n \}.
    \]
    If $n$ is a natural number, we often write $X^n$ as shorthand for the $n$-fold Cartesian product $X^n \coloneqq \prod_{1 \le i \le n} X$.
\end{df}


\begin{df}[Finite choice]{Lemma 3.5.12}
    Let $n \ge 1$ be a natural number, and for each natural number $1 \le i \le n$, let $X_i$ be a non-empty set.
    Then there exists an $n$-tuple $(x_i)_{1 \le i \le n}$ such that $x_i \in X_i$ for all $1 \le i \le n$.
    In other words, if each $X_i$ is non-empty, then the set $\prod_{1 \le i \le n} X_i$ is also non-empty.
\end{df}


\subsection{Cardinality of sets}

\begin{df}[Equal cardinality]{Definition 3.6.1}
    We say that two sets $X$ and $Y$ have \emph{equal cardinality} iff there exists a bijection $f : X \to Y$ from $X$ to $Y$.
\end{df}


\begin{df}{Proposition 3.6.4}
    Let $X$, $Y$, $Z$ be sets.
    Then $X$ has equal cardinality with $X$.
    If $X$ has equal cardinality with $Y$, then $Y$ has equal cardinality with $X$.
    If $X$ has equal cardinality with $Y$ and $Y$ has equal cardinality with $Z$, then $X$ has equal cardinality with $Z$.
\end{df}


\begin{df}{Definition 3.6.5}
    Let $n$ be a natural number.
    A set $X$ is said to have \emph{cardinality} $n$, iff it has equal cardinality with $\{ i \in \mathbb{N} : 1 \le i \le n \}$.
    We also say that $X$ \emph{has $n$ elements} iff it has cardinality $n$.
\end{df}


\begin{df}[Uniqueness of cardinality]{Proposition 3.6.8}
    Let $X$ be a set with some cardinality $n$.
    Then $X$ cannot have any other cardinality, i.e., $X$ cannot have cardinality $m$ for any $m \ne n$.
\end{df}


\begin{df}{Lemma 3.6.9}
    Suppose that $n \ge 1$, and $X$ has cardinality $n$.
    Then $X$ is non-empty, and if $x$ is any element of $X$, then the set $X - \{ x \}$ (i.e., $X$ with the element $x$ removed) has cardinality $n - 1$.
\end{df}


\begin{df}[Finite sets]{Definition 3.6.10}
    A set is \emph{finite} iff it has cardinality $N$ for some natural number $n$; otherwise, the set is called \emph{infinite}.
    If $X$ is a finite set, we use $\#(X)$ to denote the cardinality of $X$.
\end{df}


\begin{df}{Theorem 3.6.12}
    The set of natural numbers $\mathbb{N}$ is infinite.
\end{df}


\begin{df}[Cardinal arithmetic]{Proposition 3.6.14}
    \begin{enumerate}[topsep=0mm,label=(\alph*)]
        \item Let $X$ be a finite set, and let $x$ be an object which is not an element of $X$.
            Then $X \cup \{ x \}$ is finite and $\#(X \cup \{ x \}) = \#(X) + 1$.
            
        \item  Let $X$ and $Y$ be finite sets.
            Then $X \cup Y$ is finite and $\#(X \cup Y ) \le \#(X) + \#(Y).$
            If in addition X and Y are disjoint (i.e., $X \cap Y = \varnothing$), then $\#(X \cup Y) = \#(X) + \#(Y)$.

        \item Let $X$ be a finite set, and let $Y$ be a subset of $X$.
            Then $Y$ is finite, and $\#(Y) \le \#(X)$.
            If in addition $Y \ne X$ (i.e., $Y$ is a proper subset of $X$), then we have $\#(Y) < \#(X)$.

        \item If $X$ is a finite set, and $f : X \to Y$ is a function, then $f(X)$ is a finite set with $\#(f(X)) \le \#(X)$.
            If in addition $f$ is one-to-one, then $\#(f(X)) = \#(X)$.
        
        \item Let $X$ and $Y$ be finite sets.
            Then Cartesian product $X \times Y$ is finite and $\#(X \times Y) = \#(X) \times \#(Y)$.

        \item Let $X$ and $Y$ be finite sets.
            Then the set $Y^X$ (defined in Axiom 3.10) is finite and $\#(Y^X) = \#(Y)^{\#(X)}$.
    \end{enumerate}
\end{df}


\pagebreak

\section{Integers and rationals}
\subsection{The integers}

\begin{df}[Integers]{Definition 4.1.1}
    An \emph{integer} is an expression of the form $a \minus b$, where $a$ and $b$ are natural numbers.
    Two integers are considered to be equal, $a \minus b = c \minus d$, if and only if $a + d = c + b$.
    We let $\mathbb{Z}$ denote the set of all integers.
\end{df}


\begin{df}{Definition 4.1.2}
    The sum of two integers, $(a \minus b) + (c \minus d)$, is defined by the formula
    \[
        (a \minus b) + (c \minus d) \coloneqq (a + c) \minus (b + d).
    \]
    The product of two integers $(a \minus b) \times (c \minus d)$, is defined by
    {\setlength{\belowdisplayskip}{-10pt} \setlength{\belowdisplayshortskip}{-10pt}
    \[
        (a \minus b) \times (c \minus d) \coloneqq (ac + bd) \minus (ad + bc).
    \]
    }
\end{df}


\begin{df}[Negation of integers]{Definition 4.1.4}
    If $(a \minus b)$ is an integer, we define the negation $-(a \minus b)$ to be the integer $(b \minus a)$.
    In particular, if $n = n \minus 0$ is a positive natural number, we can define its negation $-n = 0 \minus n$.
\end{df}


\begin{df}[Trichotomy of integers]{Lemma 4.1.5}
    Let $x$ be an integer.
    Then exactly one of the following three statements is true:
    (a) $x$ is zero;
    (b) $x$ is equal to a positive natural number $n$; or
    (c) $x$ is the negation $-n$ of a positive natural number $n$.
\end{df}


\begin{df}[Laws of algebra for integers]{Proposition 4.1.6}
    Let $x$, $y$, $z$ be integers.
    Then we have
    {\setlength{\abovedisplayskip}{0pt}
    \begin{align*}%
        x + y &= y + x \\
        (x + y) + z &= x + (y + z) \\
        x + 0 = 0 + x &= x \\
        x + (-x) = (-x) + x &= 0 \\
        xy &= yx \\
        (xy)z &= x(yz) \\
        x1 = 1x &= x \\
        x(y + z) &= xy + xz \\
        (y + z)x &= yx + zx.%
    \end{align*}%
    }%
    \textit{Remark. } The above set of nine identities assert that the integers form a \emph{commutative ring}.
    If one deleted the identity $xy = yx$, then they would only assert that the integers form a \emph{ring}.
\end{df}


\begin{df}[Integers have no zero divisors]{Proposition 4.1.8}
    Let $a$ and $b$ be integers such that $ab = 0$.
    Then either $a = 0$ or $b = 0$ (or both).
\end{df}


\begin{df}[Cancellation law for integers]{Corollary 4.1.9}
    If $a$, $b$, $c$ are integers such that $ac = bc$ and $c$ is non-zero, then $a = b$.
\end{df}


\begin{df}[Ordering of the integers]{Definition 4.1.10}
    Let $n$ and $m$ be integers.
    We say that $n$ is \emph{greater than or equal to} $m$, 
    and write $n \ge m$ or $m \le n$,
    iff we have $n = m + a$ for some natural number $a$.
    We say that $n$ is \emph{strictly greater than} $m$,
    and write $n > m$ or $m < n$,
    iff $n \ge m$ and $n \ne m$.
\end{df}


\begin{df}[Properties of order]{Lemma 4.1.11}
    Let $a$, $b$, $c$ be integers.
    \begin{enumerate}[topsep=0mm,label=(\alph*)]
        \item $a > b$ if and only iof $a - b$ is a positive natural number.
        \item (Addition preserves order) If $a > b$, then $a + c > b + c$.
        \item (Positive multiplication preserves order) If $a > b$ and $c$ is positive, then $ac > bc$.
        \item (Negation reverses order) If $a > b$, then $-a < -b$.
        \item (Order is transitive) If $a > b$ and $b . c$, then $a > c$.
        \item (Order trichotomy) Exactly one of the statements $a > b$, $a < b$, or $a = b$ is true.
    \end{enumerate}
\end{df}


\subsection{The rationals}

\begin{df}{Definition 4.2.1}
    A \emph{rational number} is an expression of the form $a // b$, where $a$ and $b$ are integers and $b$ is non-zero; $a // 0$ is not considered to be a rational number.
    Two rational numbers are considered to be equal, $a // b = c // d$, if and only if $ad = cb$.
    The set of all rational numbers is denoted $\mathbb{Q}$.
\end{df}


\end{document}